---
title: "Sprawozdanie 3 Statystyka"
author: "Katarzyna Stasińska"
date: "2023-11"
output: pdf_document
---

```{r,echo=FALSE}
set.seed(123)
number=100
z=qnorm(0.975,0,1)
x1=qchisq(0.975,50)
x2=qchisq(0.025,50)

ramkidanych=function(){
  df<- data.frame(matrix(ncol = 4, nrow = 0))
  x <- c("dlugosc","zawiera się","lewy koniec","prawy koniec")
  colnames(df) <- x
  return(df)
}

zawieranie=function(df,int){
  if(int==1) {print(paste("(i): ","średnia dlugosc: " ,round(mean(df$dlugosc),4), "prawdopodobieństwo pokrycia: ",sum(df$`zawiera się`)/10000))}
  if(int==2) {print(paste("(ii): ","średnia dlugosc: " ,round(mean(df$dlugosc),4), "prawdopodobieństwo pokrycia: ",sum(df$`zawiera się`)/10000))}
  if(int==3) {print(paste("(iii): ","średnia dlugosc: " ,round(mean(df$dlugosc),4), "prawdopodobieństwo pokrycia: ",sum(df$`zawiera się`)/10000))}
}

wykresy=function(nazwa,value,df){
  l=min(df$`lewy koniec`[1:number])
  r=max(df$`prawy koniec`[1:number])
  l=l-abs(0.1*l)
  r=r+abs(0.1*r)
  x=max(abs(value-l),abs(r-value))
  plot(c(value-x,value+x),c(1,number), type="n", xlab=nazwa, ylab=paste(number," przedziałów ufności"))
  segments(df$`lewy koniec`,1:number,df$`prawy koniec`,1:number, col="blue", lwd=1)
  abline(v=value,col="red",lwd=1)
}

f2=function(X,mi,sd,n){
  flaga=0
  l=mean(X)-z*sd/(n)^(0.5)
  r=mean(X)+z*sd/(n)^(0.5)
  dlugosc=r-l
  if (mi>l && mi<r){flaga=1}
  return(c(dlugosc,flaga,l,r)) 
}

f4=function(X,mi,sd,n){
  flaga=0
  sd=(sum((X-mean(X))^2)/(n-1))^(1/2) ## nieznana wariancja
  l=mean(X)-z*sd/(n)^(0.5)
  r=mean(X)+z*sd/(n)^(0.5)
  dlugosc=r-l
  if (mi>l && mi<r){flaga=1}
  return(c(dlugosc,flaga,l,r)) 
}

f6=function(X,mi,sd,n){
  flaga=0
  s2=1/n*sum((X-mi)^2)
  l=n*s2/x1
  r=n*s2/x2
  dlugosc=r-l
  if (sd^2>l && sd^2<r){flaga=1}
  return(c(dlugosc,flaga,l,r)) 
}

f8=function(X,mi,sd,n){
  flaga=0
  s2=1/n*sum((X-mean(X))^2)
  l=n*s2/x1
  r=n*s2/x2
  dlugosc=r-l
  if (sd^2>l && sd^2<r){flaga=1}
  return(c(dlugosc,flaga,l,r)) 
}

f10=function(X,p,n){
  flaga=0
  pval=sum(X>0)/n
  var=sqrt(pval*(1-pval)/n)
  l=pval-z*var
  r=pval+z*var
  dlugosc=r-l
  if (p>l && p<r){flaga=1}
  return(c(dlugosc,flaga,l,r)) 
}

eksperyment=function(rozklad,funkcja,p11,p12,p21,p22,p31,p32,mi1,mi2,mi3,sd1,sd2,sd3,n,flaga){
  df1 <- ramkidanych()
  df2 <- ramkidanych()
  df3 <- ramkidanych()
  
  for (i in 1:10000){
    if(p12 == -100){
      X1=rozklad(n,p11)
      X2=rozklad(n,p21)
      X3=rozklad(n,p31)
    }else{
      X1=rozklad(n,p11,p12)
      X2=rozklad(n,p21,p22)
      X3=rozklad(n,p31,p32)
    }
    df1[nrow(df1)+1,] <- funkcja(X1,mi1,sd1,n)
    df2[nrow(df2)+1,] <- funkcja(X2,mi2,sd2,n)
    df3[nrow(df3)+1,] <- funkcja(X3,mi3,sd3,n)
  }
  
  zawieranie(df1,1)
  zawieranie(df2,2)
  zawieranie(df3,3)
  
  if(flaga==0){
    par(mfrow=c(1, 3))
    wykresy('średnia',mi1,df1)
    wykresy('średnia',mi2,df2)
    wykresy('średnia',mi3,df3)
  }
  if(flaga==1){
    par(mfrow=c(1, 3))
    wykresy('wariancja',sd1,df1)
    wykresy('wariancja',sd2,df2)
    wykresy('wariancja',sd3,df3)
  }
}

```

# Zadanie 1 

NOTATKI BRUDNOPIS

# Zadanie 2

### wyniki eksperymentu a)

```{r, echo=FALSE}
eksperyment(rnorm,f2,0,1,0,2,0,3,0,0,0,1,2,3,50,0)
```

### wyniki eksperymentu b)

```{r, echo=FALSE}
eksperyment(rlogis,f2,0,1,0,2,0,3,0,0,0,sqrt(1^2*pi^2/3),sqrt(2^2*pi^2/3),sqrt(3^2*pi^2/3),50,0)
```

### wyniki eksperymentu c)

```{r, echo=FALSE}
n=50
X1=rcauchy(50,0,1)
X2=rcauchy(50,0,2)
X3=rcauchy(50,0,3)
sd1=(sum((X1-mean(X1))^2)/(n-1))^(1/2)
sd2=(sum((X2-mean(X2))^2)/(n-1))^(1/2)
sd3=(sum((X3-mean(X3))^2)/(n-1))^(1/2)
eksperyment(rcauchy,f2,0,1,0,2,0,3,0,0,0,sd1,sd2,sd3,50,0)
```

### wyniki eksperymentu d)

```{r, echo=FALSE}
eksperyment(rexp,f2,1,-100,1/2,-100,1/3,-100,1,2,3,1,2,3,50,0)
```

### wyniki eksperymentu e)

```{r, echo=FALSE}
eksperyment(rchisq,f2,1,-100,2,-100,3,-100,1,2,3,2^(1/2),4^(1/2),6^(1/2),50,0)
```

# Zadanie 4

### wyniki eksperymentu a)

```{r, echo=FALSE}
eksperyment(rnorm,f4,0,1,0,2,0,3,0,0,0,1,2,3,50,0)
```

### wyniki eksperymentu b)

```{r, echo=FALSE}
eksperyment(rlogis,f4,0,1,0,2,0,3,0,0,0,sqrt(1^2*pi^2/3),sqrt(2^2*pi^2/3),sqrt(3^2*pi^2/3),50,0)
```

### wyniki eksperymentu c)

```{r, echo=FALSE}
n=50
X1=rcauchy(50,0,1)
X2=rcauchy(50,0,2)
X3=rcauchy(50,0,3)
sd1=(sum((X1-mean(X1))^2)/(n-1))^(1/2)
sd2=(sum((X2-mean(X2))^2)/(n-1))^(1/2)
sd3=(sum((X3-mean(X3))^2)/(n-1))^(1/2)
eksperyment(rcauchy,f4,0,1,0,2,0,3,0,0,0,sd1,sd2,sd3,50,0)
```

### wyniki eksperymentu d)

```{r, echo=FALSE}
eksperyment(rexp,f4,1,-100,1/2,-100,1/3,-100,1,2,3,1,2,3,50,0)
```

### wyniki eksperymentu e)

```{r, echo=FALSE}
eksperyment(rchisq,f4,1,-100,2,-100,3,-100,1,2,3,2^(1/2),4^(1/2),6^(1/2),50,0)
```

# Zadanie 6

### wyniki eksperymentu a)

```{r, echo=FALSE}
eksperyment(rnorm,f6,0,1,0,2,0,3,0,0,0,1,2,3,50,1)
```

### wyniki eksperymentu b)

```{r, echo=FALSE}
eksperyment(rlogis,f6,0,1,0,2,0,3,0,0,0,sqrt(1^2*pi^2/3),sqrt(2^2*pi^2/3),sqrt(3^2*pi^2/3),50,1)
```

### wyniki eksperymentu c)

```{r, echo=FALSE}
n=50
X1=rcauchy(50,0,1)
X2=rcauchy(50,0,2)
X3=rcauchy(50,0,3)
sd1=(sum((X1-mean(X1))^2)/(n-1))^(1/2)
sd2=(sum((X2-mean(X2))^2)/(n-1))^(1/2)
sd3=(sum((X3-mean(X3))^2)/(n-1))^(1/2)
eksperyment(rcauchy,f6,0,1,0,2,0,3,0,0,0,sd1,sd2,sd3,50,1)
```

### wyniki eksperymentu d)

```{r, echo=FALSE}
eksperyment(rexp,f6,1,-100,1/2,-100,1/3,-100,1,2,3,1,2,3,50,1)
```

### wyniki eksperymentu e)

```{r, echo=FALSE}
eksperyment(rchisq,f6,1,-100,2,-100,3,-100,1,2,3,2^(1/2),4^(1/2),6^(1/2),50,1)
```

# Zadanie 8

### wyniki eksperymentu a)

```{r, echo=FALSE}
eksperyment(rnorm,f8,0,1,0,2,0,3,0,0,0,1,2,3,50,1)
```

### wyniki eksperymentu b)

```{r, echo=FALSE}
eksperyment(rlogis,f8,0,1,0,2,0,3,0,0,0,sqrt(1^2*pi^2/3),sqrt(2^2*pi^2/3),sqrt(3^2*pi^2/3),50,1)
```

### wyniki eksperymentu c)

```{r, echo=FALSE}
n=50
X1=rcauchy(50,0,1)
X2=rcauchy(50,0,2)
X3=rcauchy(50,0,3)
sd1=(sum((X1-mean(X1))^2)/(n-1))^(1/2)
sd2=(sum((X2-mean(X2))^2)/(n-1))^(1/2)
sd3=(sum((X3-mean(X3))^2)/(n-1))^(1/2)
eksperyment(rcauchy,f8,0,1,0,2,0,3,0,0,0,sd1,sd2,sd3,50,1)
```

### wyniki eksperymentu d)

```{r, echo=FALSE}
eksperyment(rexp,f8,1,-100,1/2,-100,1/3,-100,1,2,3,1,2,3,50,1)
```

### wyniki eksperymentu e)

```{r, echo=FALSE}
eksperyment(rchisq,f8,1,-100,2,-100,3,-100,1,2,3,2^(1/2),4^(1/2),6^(1/2),50,1)
```